### 2005-08-30 docs/paper/kcorrect_tables.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@499 75dff457-8751-0410-ab1a-cd701654ce6b

\begin{deluxetable}{rrrrr}
\tablewidth{0pt}
\tablecolumns{5}
\tablecaption{\label{solarmagnitudes} Properties of various filters}
\tablehead{ Band & $\lambda_{\mathrm{eff}}$ (\AA) & $m_{\mathrm{AB}} -
	m_{\mathrm{Vega}}$ & $M_\odot$ (AB) &
	$M_\odot$ (Vega)}
\startdata
\input{solarmagnitudes.tex}
\enddata
\tablecomments{Uses model solar spectrum and model Vega spectrum of
	\citet{kurucz91a}. Effective wavelength is defined in the text.}
\end{deluxetable}

\begin{deluxetable}{cc}
\tablewidth{0pt}
\tablecolumns{2}
\tablecaption{\label{lineareqs} Conversions among various filters}
\tablehead{ Equation & Color dispersion }
\startdata
\input{lineareqs.tex}
\enddata
\tablecomments{Uses AB magnitudes throughout.}
\end{deluxetable}