### 2005-09-07 docs/paper/kcorrect_tables.tex

\begin{deluxetable}{rrrrr}
\tablewidth{0pt}
\tablecolumns{5}
\tablecaption{\label{solarmagnitudes} Properties of various filters}
\tablehead{ Band & $\lambda_{\mathrm{eff}}$ (\AA) & $m_{\mathrm{AB}} -
	m_{\mathrm{Vega}}$ & $M_\odot$ (AB) &
	$M_\odot$ (Vega)}
\startdata
\input{solarmagnitudes.tex}
\enddata
\tablecomments{Uses model solar spectrum and model Vega spectrum of
	\citet{kurucz91a}. Effective wavelength is defined in the text.}
\end{deluxetable}

\begin{deluxetable}{lr}
\tablewidth{0pt}
\tablecolumns{2}
\tablecaption{\label{lineareqs} Conversions among various filters}
\tablehead{ Equation & Color dispersion }
\startdata
\input{lineareqs.tex}
\enddata
\tablecomments{Uses AB magnitudes throughout.}
\end{deluxetable}

\begin{deluxetable}{llcl}
\tablewidth{0pt}
\tablecolumns{4}
\tablecaption{\label{derived} HDU of {\tt k\_nmf\_derived.default.fits}}
\tablehead{ Number & Name & Dimensions & Description}
\startdata
0 & {\tt templates} & $N_{\mathrm{basis}} \times N_t$ & Coefficients
in basis space for each template \cr
1 & {\tt spec} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template \cr
2 & {\tt spec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template, lines removed \cr
3 & {\tt spec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template, without dust \cr
4 & {\tt spec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template, without lines or dust \cr
5 & {\tt rawspec} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
each template \cr
6 & {\tt rawspec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
each template, lines removed \cr
7 & {\tt rawspec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
each template, without dust \cr
8 & {\tt rawspec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ &
Unsmoothed spectrum of each template, without lines or dust \cr
9 & {\tt lspec} & $N_{\mathrm{spec}} \times N_t$ & emission line spectrum \cr
10 & {\tt extinction} & $N_{\mathrm{spec}} \times N_t$ & extinction as
a function of wavelength \cr
11 & {\tt lambda} & $N_{\mathrm{spec}}$ & Wavelength grid (\AA) \cr
12 & {\tt sfr} & $N_{\mathrm{age}} \times N_t$ & Star-formation rate
($M_\odot$ per year) as a function of time \cr
13 & {\tt metallicity} & $N_{\mathrm{age}} \times N_t$ & Average
metallicity as a function of time \cr
14 & {\tt ages} & $N_{\mathrm{age}}$ & Age grid (yrs) \cr
15 & {\tt dage} & $N_{\mathrm{age}}$ & Age bin size (yrs);
multiply {\tt sfr} for total star-formation in each bin \cr
16 & {\tt mass} & $N_t$ & total mass formed in each template \cr
17 & {\tt mremain} & $N_t$ & total current stellar mass in each template \cr
18 & {\tt mets} & $N_t$ & average metallicity of current stars in each
template \cr
19 & {\tt m300} & $N_t$ & total stellar mass formed in last 300 Myrs in each template \cr
20 & {\tt m1000} & $N_t$ & total stellar mass formed in last 1 Gyr in each template \cr
21 & {\tt basis\_ages} &
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
ages of each stellar population basis vector  \cr
22 & {\tt basis\_mets} & 
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
metallicity of each stellar population basis vector  \cr
23 & {\tt basis\_dusts} & 
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
dust properties of each stellar population basis vector \cr
24 & {\tt basis\_mremain} & 
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
fraction of original stellar mass surviving for age of each basis vector  \cr
\enddata 
\tablecomments{All quantities are in floating point (four byte)
precision, except {\tt dusts} which is an array of structures, whose
format is described in text.}
\end{deluxetable}