### 2006-02-16 docs/paper/kcorrect_tables.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@549 75dff457-8751-0410-ab1a-cd701654ce6b

\begin{deluxetable}{rrrrr}
\tablewidth{0pt}
\tablecolumns{5}
\tablecaption{\label{solarmagnitudes} Properties of various filters}
\tablehead{ Band & $\lambda_{\mathrm{eff}}$ (\AA) & $m_{\mathrm{AB}} -
	m_{\mathrm{Vega}}$ & $M_\odot$ (AB) &
	$M_\odot$ (Vega)}
\startdata
\input{solarmagnitudes.tex}
\enddata
\tablecomments{Uses model solar spectrum and model Vega spectrum of
	\citet{kurucz91a}. Effective wavelength is defined in the text. The
	$UBRVI$ filters are those of \citet{bessell90a}. The $ugriz$ filters
	are those determined by Mamoru Doi, Daniel Eisenstein, and James
	Gunn and available on the SDSS DR4 web site.\footnote{{\tt
	http://www.sdss.org/dr4/}} The $JHK_s$ filters are those from
	\citet{cohen03a}. }
\end{deluxetable}

\begin{deluxetable}{lr}
\tablewidth{0pt}
\tablecolumns{2}
\tablecaption{\label{lineareqs} Conversions among various filters}
\tablehead{ Equation & Color dispersion }
\startdata
\input{lineareqs.tex}
\enddata
\tablecomments{Uses AB magnitudes throughout.}
\end{deluxetable}

\begin{deluxetable}{llcl}
\tablewidth{0pt}
\tablecolumns{4}
\tablecaption{\label{derived} Header and Data Units (HDUs) of {\tt k\_nmf\_derived.default.fits}}
\tablehead{ Number & Name & Dimensions & Description}
\startdata
0 & {\tt templates} & $N_{\mathrm{basis}} \times N_t$ & Coefficients
in basis space for each template \cr 1 & {\tt spec} &
$N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of each template
\cr 2 & {\tt spec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Smoothed
spectrum of each template, lines removed \cr 3 & {\tt spec\_nd} &
$N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of each template,
without dust \cr 4 & {\tt spec\_nl\_nd} & $N_{\mathrm{spec}} \times
N_t$ & Smoothed spectrum of each template, without lines or dust \cr 5
& {\tt rawspec} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum
of each template \cr 6 & {\tt rawspec\_nl} & $N_{\mathrm{spec}} \times
N_t$ & Unsmoothed spectrum of each template, lines removed \cr 7 &
{\tt rawspec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed
spectrum of each template, without dust \cr 8 & {\tt rawspec\_nl\_nd}
& $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of each
template, without lines or dust \cr 9 & {\tt lspec} &
$N_{\mathrm{spec}} \times N_t$ & emission line spectrum \cr 10 & {\tt
extinction} & $N_{\mathrm{spec}} \times N_t$ & extinction as a
function of wavelength \cr 11 & {\tt lambda} & $N_{\mathrm{spec}}$ &
Wavelength grid (\AA) \cr 12 & {\tt sfr} & $N_{\mathrm{age}} \times
N_t$ & Star-formation rate ($M_\odot$ per year) as a function of time
\cr 13 & {\tt metallicity} & $N_{\mathrm{age}} \times N_t$ & Average
metallicity as a function of time \cr 14 & {\tt ages} &
$N_{\mathrm{age}}$ & Age grid (yrs) \cr 15 & {\tt dage} &
$N_{\mathrm{age}}$ & Age bin size (yrs); {\tt sfr}$\times${\tt dage}
$=$ total star-formation in each bin \cr 16 & {\tt mass} & $N_t$ &
total mass formed in each template \cr 17 & {\tt mremain} & $N_t$ &
total current stellar mass in each template \cr 18 & {\tt mets} &
$N_t$ & average metallicity of current stars in each template \cr 19 &
{\tt m300} & $N_t$ & total stellar mass formed in last 300 Myrs in
each template \cr 20 & {\tt m1000} & $N_t$ & total stellar mass formed
in last 1 Gyr in each template \cr 21 & {\tt basis\_ages} &
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
ages of each stellar population basis vector \cr 22 & {\tt
basis\_mets} & $N_{\mathrm{age}} \times N_{\mathrm{mets}} \times
N_{\mathrm{dust}}$ & metallicity of each stellar population basis
vector \cr 23 & {\tt basis\_dusts} & $N_{\mathrm{age}} \times
N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ & dust properties of each
stellar population basis vector \cr 24 & {\tt basis\_mremain} &
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
fraction of original stellar mass surviving for each basis vector \cr
\enddata 
\tablecomments{All quantities are in floating point (four byte)
precision, except {\tt dusts} which is an array of structures, whose
format is described in text.}
\end{deluxetable}